Collective excitations in competing phases in two and three dimensions

### 2023-12-07 paper/main.tex

We observe that the entire Hamiltonian and thus all expectation values only depend on
\begin{equation}
    \hat{\gamma}(\vk) \coloneqq \frac{1}{D} \sum_{\alpha=1}^D \cos(k_\alpha)\,,
\end{equation}
i.e., for any operator $\hat{O}$ the following relation holds 
\begin{equation}
    \label{eqn:equal_expecs}
    \langle \hat{O}_{\vk} \rangle = \langle \hat{O}_{\vk'} \rangle \eqqcolon \langle \hat{O}_{\gamma} \rangle\,,
\end{equation}
if $\hat{\gamma}(\vk) = \hat{\gamma}(\vk')$.
While two-dimensional systems around the size of $100\times100$ lattice sites can still be solved by using momentum sums,
computing solutions for large three-dimensional systems becomes impossible due to the $N=L^3$ scaling.
This issue can be resolved by using the aforementioned fact and replacing the momentum sums with energy integrals using the bare density of states (DOS)
\begin{equation}
    \rho(\gamma) \coloneqq  \frac{1}{N} \sum_{\vk} \delta \left(\gamma - \hat{\gamma} (\vk) \right)\,.
\end{equation}
As an example, we consider 
\begin{equation}
    \Delta_\text{SC} = \frac{U}{2N} \int_{-1}^{1} \dgamma \rho(\gamma) \langle f_{\gamma} \rangle\,.
\end{equation}
We will be using the exact densities of states for the square and the simple cubic lattice found in Ref. \cite{Hanisch97}.
This allows us to access both two- and three-dimensional models on equal footing.
We can compute any phase as long as it does not introduce another kind of momentum dependence.

@@

We start with some operator set $\mathcal{B}$ that is complete with respect to commutation with a Hamiltonian $H$, 
i.e., any commutator $[H, A]$ for any $A \in \mathcal{B}$ can be represented by linear combinations of operators in $\mathcal{B}$.
We can then express any time-depend operator from this set as
\begin{equation}
    a(t) = \sum_j c_j(t) A_j\,,
\end{equation}
where $c_j(t)$ captures the entire time dependence. 
Inserting this into the Heisenberg equation of motion and applying an operator scalar product $(A_i|\cdot)$ to both sides of the equation yields
\begin{equation}
    \label{eqn:heisenberg}
    \begin{aligned}
        \ddt a(t) = \sum_j \ddt c_j(t) A_j &= i \sum_j c_j(t) [H, A_j] \\
        \Rightarrow \sum_j \underbrace{(A_i | A_j)}_{\coloneqq \mN_{ij}} \ddt c_j(t) &= i \sum_j \underbrace{(A_i | [H, A_j])}_{\coloneqq \mM_{ij}} c_j(t) \\
        \Rightarrow \mN \ddt \vec{c}(t) &= i \mM \vec{c}(t)\,.
    \end{aligned}
\end{equation}
The matrices $\mN$ and $\mM$ contain all the dynamics and energetic properties of the system.
The advantage is that one can now handle simple matrices rather than operators acting on an enormous Hilbert space.
The size of these matrices depends entirely on the set of operators.

@@

To study the collective excitation, we investigate various retarded Green's functions 
\begin{equation}
    G_{AB}^\text{ret} (t) = - i \langle [A(t), B] \rangle \Theta(t)\,,
\end{equation}
where $\Theta(t)$ is the Heaviside function, and their Fourier-transformed
\begin{equation}
    \label{eqn:standard_gf}
    G_{AB}(\omega + i0^+) = -i \int_0^{\infty} e^{izt} \langle [A(t), B] \rangle \mathrm{d}t\,.
\end{equation}
Our previous choice of pseudo scalar product enables us to write down a matrix-valued Green's functions in terms of the aforementioned matrices
\begin{align}
    \label{eqn:green_function}
    \mathcal{G}(z &= \omega + i0^+) = \mN \frac{1}{-z \mN - \mM} \mN \\
        &\coloneqq  -\mN \mathcal{R}(z) \mN\,,
\end{align}
where we used the resolvent
\begin{equation}
    \label{eqn:resolvent}
    \mathcal{R}(z) \coloneqq  \frac{1}{z \mN + \mM}\,.
\end{equation}
The entries of this matrix $\mathcal{G}(z)$ are the Fourier transformed Green's functions mentioned in \eqref{eqn:standard_gf} with respect to the operators from the initially chosen set.
For instance, assume the operator in the set is $f_{\vk}$, then the top left element of $\mathcal{G}$ is the Green's function 
\begin{equation}
    \mathcal{G}_{00}(z = \omega +i0^+) =  G(z) = -i \int_0^{\infty} \langle [f_{\vk}(t), f_{\vk}^\dagger(0)] \rangle e^{izt} \mathrm{d}t\,.
\end{equation}
For a detailed derivation of this statement, we refer the reader to the appendix \textcolor{red}{!REF!}.

Our operator set consists of operators of the type
\begin{equation}
    A_\gamma \coloneqq  \frac{1}{\sqrt{N}} \sum_{\vk} \delta (\gamma - \hat{\gamma}( \vk )) A_{\vk}\,,
\end{equation}
where $A_{\vk}$ represents each type of operator in \eqref{eqn:operators} and their Hermitian conjugates.
Different indices are taken as additional set operators, e.g. $n_{\vk \up}$, and $n_{\vk \down}$ are taken into the set separately.
We then order these operators in a way reminiscent of the $x$- and $p$-terms in the harmonic oscillator, i.e.
\begin{equation}
    X_i \coloneqq  A_i + A_i^\dagger\,,\quad P_i \coloneqq  A_i - A_i^\dagger\,.
\end{equation}
Our set is then $\mathcal{B}_{XP} \coloneqq  \{ X_i, P_i \}$.
If some operators would be 0 ($P_i$ for any $n$-type term) or duplicates (certain $X_i$ for $g$-type terms), we omit them.

Due to this choice of set, the matrices occurring in \eqref{eqn:heisenberg} have a special block structure
\begin{equation}
    \label{eqn:xp_set}
    \mM = \begin{pmatrix}
        \mathcal{K}_+ & \kappa \\ \kappa^\dagger & \mathcal{K}_-
    \end{pmatrix}\,,\quad \mN = \begin{pmatrix}
        \Lambda_+ & \mathcal{L} \\ \mathcal{L}^\dagger & \Lambda_-
    \end{pmatrix}\,,
\end{equation}
where $\Im [\mathcal{K}_\pm] = \Re [\kappa] = 0$ and $\Re [\Lambda_\pm] = \Im [\mathcal{L}] = 0$.
Furthermore, if all occurring expectation values are real, both matrices have large empty blocks.
This condition is fulfilled for all cases investigated in this article.
Additionally, we can utilize this block structure to rewrite \eqref{eqn:resolvent} to
\begin{equation}
    \mathcal{R}(z) \vert_X = - \check{N}^{-1/2} \frac{1}{z^2 - \underbrace{\check{N}^{-1/2} \mathcal{K}_+ \check{N}^{-1/2}}_{\coloneqq  \check{M}}} \check{N}^{-1/2}\,,
\end{equation}
where $\vert_X$ denotes the block corresponding to the $X$-terms in the set and $\check{N} \coloneqq  \mathcal{L} \mathcal{K}_-^{-1} \mathcal{L}^\dagger$.
An analogous form can be found for the $P$-block.

@@

\begin{widetext}
\begin{equation}
    \left( z^2 - \begin{pmatrix}
        a_0 & b_1 & 0 & 0 & \cdots \\
        b_1 & a_1 & b_2 & 0 & \cdots \\
        0 & b_2 & a_2 & b_3 & \cdots \\
        \vdots & \vdots & \vdots & \vdots & \ddots
    \end{pmatrix} \right)_{00}^{-1} = \dfrac{1}{z^2 - a_0 - \dfrac{b_1^2}{z^2 - a_1 - \dfrac{b_2^2}{ z^2 - a_2 - \hdots}}}\,\,
\end{equation}
\end{widetext}

where $a_i$ and $b_i$ are the Lanczos coefficients \cite{PettiforRecursion,ViswanathRecursion}.
In the single-band case, the coefficients approach the limit
\begin{equation}
    \label{eqn:inf_lanczos}
    a_\infty = \frac{\omega_+ + \omega_-}{2}\text{  and  } b_\infty = \frac{\omega_+ - \omega_-}{4}\,,
\end{equation}
where $\omega_\pm$ represents the upper and lower band edge, respectively.
This behaviour allows us to terminate the continued fraction, if the convergence is good enough, using the so-called square root terminator
\begin{equation}
    T(\omega) = \frac{1}{2b_\infty^2} \left( \omega - a_\infty \mp \sqrt{(\omega - a_\infty)^2 - 4 b_\infty^2} \right)\,,
\end{equation}
where the negative sign is to be chosen for $\omega - a_\infty > 2b_\infty$ and $|\omega - a_\infty| < 2b_\infty$ and the positive one otherwise.

@@

In this section, we will be investigating four different diagonal Green's functions $G_{AA}(\omega + i0^+)$, namely with
\begin{subequations}
    \label{eqn:resolvent_bases}
    \begin{align}
        A_\text{Higgs} &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( f_{\vk} + f_{\vk}^\dagger \right)\,,\\
        A_\text{Phase} &= \frac{i}{\sqrt{N}} \sum_{\vk} \left( f_{\vk} - f_{\vk}^\dagger \right)\,,\\
        A_\text{CDW} &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( g_{\vk \up} + g_{\vk \down} \right)\,,\\
        A_\text{AFM} &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( g_{\vk \up} - g_{\vk \down} \right)\,,
    \end{align}
\end{subequations}
where $N$ is the number of lattice sites in the system. 
Note, that this choice of normalization cancels out with other system size-dependent terms so that our final results do not depend on it at all.
The first operator catches the amplitude mode of the $s$-wave superconducting state, while the second one catches the phase mode \cite{Fan22}.
The remaining two capture the collective behaviour of the CDW and AFM states, respectively.

@@

In this appendix section, we will derive the relation between the matrices $\mM, \mN$ from \eqref{eqn:heisenberg} and the Green's functions of the system.
For that, consider their definition
\begin{align}
    G_{AB^\dagger}^\text{ret} (t) &= - i \langle [ A(t), B^\dagger(0) ] \rangle \theta(t) \nonumber \\ 
    \Rightarrow G^\text{ret} (z = \omega + i0^+) &= - i \int_0^\infty \exp(izt) \langle [ A(t), B^\dagger(0) ] \rangle \mathrm{d}t \nonumber
\end{align}
Since $A$ and $B$ are included in the operator set, we can rewrite them as dot-products of the vector
\begin{equation}
    \vec{A} \coloneqq  \left( A_1, A_2, \dots, A_N \right)\,, \quad
    \vec{A}^\dagger \coloneqq  \begin{pmatrix}
        A_1^\dagger \\ A_2^\dagger \\ \vdots \\ A_N^\dagger
    \end{pmatrix}\,,
\end{equation}
and some other vector with complex-valued vector $\vec{a}$ which has its coefficients set in such a way,
that $\vec{a} \cdot \vec{A}$ is equal to the desired operator. 
For instance, $\vec{a} \equiv \vec{e}_1$ in order to represent $A_1$.
Furthermore, assuming, that $\mN$ is invertible, we can easily solve the differential equation \eqref{eqn:heisenberg} to obtain
\begin{equation}
    \vec{c}(t) = \exp \left( i \mN^{-1} \mM t \right) \vec{c}(0)\,.
\end{equation}
This allows us to finally obtain
\begin{align}
    G^\text{ret} (z) &= i \int_0^\infty \exp(izt) \langle [ (\vec{A} \vec{b}(0))^\dagger, \vec{A} \vec{a}(t) ] \rangle\ \mathrm{d}t \nonumber \\
        &= i \int_0^\infty \exp(izt) \vec{b}^\dagger(0) \underbrace{\langle [ \vec{A}^\dagger, \vec{A} ] \rangle}_{\equiv \mN} \vec{a}(t) \mathrm{d}t \nonumber \\
        &= i  \int_0^\infty \vec{b}^\dagger(0) \mN \exp \left( i \left(\mN^{-1} \mM + z \right) t \right) \vec{a}(0) \mathrm{d}t  \nonumber \\
        &= - \vec{b}^\dagger(0) \left[ \mN \frac{1}{\mN^{-1} \mM + z} \right] \vec{a}(0) \nonumber \\
        &= - \vec{b}^\dagger(0) \left[ \mN \frac{1}{ \underbrace{\mM + z \mN}}_{\equiv \mathcal{R}(z)} \mN \right] \vec{a}(0)\,,
\end{align}
where $\vec{a}(0)$ and $\vec{b}(0)$ are chosen such, that their product with $\vec{A}$ yields $A(0)$ and $B(0)$, respectively.
The task at hand is finding the resolvent as $\mathcal{R}$.
Normally, one would desire to compute
\begin{align}
    \mathcal{R}(z) = \mN^{-1/2} \frac{1}{z +\mN^{-1/2} \mM \mN^{-1/2}} \mN^{-1/2}\,,
\end{align}
and then utilize a Lanczos algorithm to obtain a tridiagonal form of $\mN^{-1/2} \mM \mN^{-1/2}$.
However, $\mN$ is not positive in our case due to the nature of the commutator, which means that $\mN$ cannot have a real square root.
One could also fathom computing
\begin{align}
    \mathcal{R}(z) = \frac{1}{z} \mM^{-1/2} \frac{1}{\mM^{-1/2}\mathcal{N}\mM^{-1/2} + \frac{1}{z} } \mM^{-1/2}\,,
\end{align}
which does in fact work, if $\mM$ is positive, however resolving feature that lie near $z=0$ are difficult to resolve.
Nevertheless, let us show that $\mathcal{M}$ is positive semidefinite if the system is in thermal equilibrium.
For any $\vec{x} \in \mathbb{C}^N$, consider
\begin{align}
    \vec{x}^\dagger \mM \vec{x} &= \sum_{ij} x_i^* x_j \mM_{ij} \nonumber \\
        &= \langle \left[ \sum_i x_i^* A_i^\dagger, \left[ H, \sum_j x_j A_j \right] \right]  \rangle \nonumber \\
        &= \langle [B^\dagger, [H, B]] \rangle\,,\quad B \coloneqq  \sum_j x_j A_j\,.
\end{align}
Expanding the thermal expectation value $\langle O \rangle = \frac{1}{Z} \Trace [\exp(-\beta H_\text{MF}) O]$, where $\beta$ is the inverse temperature and $Z$ the partition function, and the commutators, we obtain

@@

We embrace the matrix structure in \eqref{eqn:xp_set} and assume that all matrix entries are real.
Then, with $m|_X$ denoting the block of a matrix $m$ spanned by the $X$-operators in the set, we can calculate
\begin{align}
    r(z) &\coloneqq  \mathcal{R}(z)|_X = \left. \frac{1}{\mM + z \mN} \right\vert_X \nonumber \\
        &= \left[ \frac{1}{\mM} - \frac{1}{\mM} z \mN \frac{1}{\mM} + \frac{1}{\mM} z \mN \frac{1}{\mM} z \mN \frac{1}{\mM} - \cdots \right]_X \nonumber \\
        &= \left. \frac{1}{\mM} \sum_{j=0}^\infty \left( -z \mN \frac{1}{\mM} \right)^j \right\vert_X \,.
\end{align}
We may omit every second element of the sum, as we want to obtain the upper left block of the matrix, and $\mN$ always swaps the upper left and lower right block.
Thus, we obtain
\begin{align}
    r(z) &= \left. \frac{1}{\mM} \sum_{j=0}^\infty z^{2j} \left( \mN \frac{1}{\mM} \right)^{2j} \right\vert_X \nonumber \\
        &= \left. \frac{1}{\mM - z^2 \mN \mM^{-1} \mN} \right\vert_X \nonumber \\
        &= \frac{1}{\mathcal{K}_+ - z^2 \mathcal{L} \mathcal{K}_-^{-1} \mathcal{L}^\dagger}\,.
\end{align}
Next, we define $\mathcal{L} \mathcal{K}_-^{-1} \mathcal{L}^\dagger \coloneqq  \check{N}$ for brevity.
Since $\mathcal{M}$ is positive semidefinite and block diagonal, its blocks $\mathcal{K}_\pm$ must be positive semidefinite as well.
Therefore, $\check{N}$ is also positive semidefinite and we can define its square root.
The inverse in the following can be replaced by the pseudo-inverse in case $\check{N}$ has a singular part.
We obtain
\begin{equation}
    r(z) = \check{N}^{-1/2} \frac{1}{\check{N}^{-1/2} \mathcal{K}_+ \check{N}^{-1/2} - z^2} \check{N}^{-1/2}\,.
\end{equation}
The matrix $\check{M} \coloneqq  \check{N}^{-1/2} \mathcal{K}_+ \check{N}^{-1/2}$ is positive semidefinite and Hermitian by the previous arguments.
Hence, we can tridiagonalize it using a standard Lanczos algorithm in order to obtain a continued fraction expansion.

>>>>

We observe that the entire Hamiltonian and thus all expectation values only depend on
\begin{equation}
    \hat{\gamma}(\vk) \coloneqq \frac{1}{D} \sum_{\alpha=1}^D \cos(k_\alpha)\,,
\end{equation}
i.e., for any operator $\hat{O}$ the following relation holds 
\begin{equation}
    \label{eqn:equal_expecs}
    \langle \hat{O}_{\vk} \rangle = \langle \hat{O}_{\vk'} \rangle \eqqcolon \langle \hat{O}( \gamma ) \rangle\,,
\end{equation}
if $\hat{\gamma}(\vk) = \hat{\gamma}(\vk')$.
While two-dimensional systems around the size of $100\times100$ lattice sites can still be solved by using momentum sums,
computing solutions for large three-dimensional systems becomes impossible due to the $N=L^3$ scaling.
This issue can be resolved by using the aforementioned fact and replacing the momentum sums with energy integrals using the bare density of states (DOS)
\begin{equation}
    \rho(\gamma) \coloneqq  \frac{1}{N} \sum_{\vk} \delta \left(\gamma - \hat{\gamma} (\vk) \right)\,.
\end{equation}
As an example, we consider 
\begin{equation}
    \Delta_\text{SC} = \frac{U}{2N} \int_{-1}^{1} \dgamma \rho(\gamma) \langle f( \gamma ) \rangle\,.
\end{equation}
We will be using the exact densities of states for the square and the simple cubic lattice found in Ref. \cite{Hanisch97}.
This allows us to access both two- and three-dimensional models on equal footing.
We can compute any phase as long as it does not introduce another kind of momentum dependence.

@@

We start with some operator set $\mathcal{B}$ that is complete with respect to commutation with a Hamiltonian $H$, 
i.e., any commutator $[H, A]$ for any $A \in \mathcal{B}$ can be represented by linear combinations of operators in $\mathcal{B}$.
We can then express any time-dependent operator from this set as
\begin{equation}
    a(t) = \sum_j c_j(t) A_j\,,
\end{equation}
where $c_j(t)$ captures the entire time dependence. 
Inserting this into the Heisenberg equation of motion and applying an operator scalar product $(A_i|\cdot)$ to both sides of the equation yields
\begin{equation}
    \label{eqn:heisenberg}
    \begin{aligned}
        \ddt a(t) = \sum_j \ddt c_j(t) A_j &= i \sum_j c_j(t) [H, A_j] \\
        \Rightarrow \sum_j \underbrace{(A_i | A_j)}_{\coloneqq \mN_{ij}} \ddt c_j(t) &= i \sum_j \underbrace{(A_i | [H, A_j])}_{\coloneqq \mM_{ij}} c_j(t) \\
        \Rightarrow \mN \ddt \vec{c}(t) &= i \mM \vec{c}(t)\,.
    \end{aligned}
\end{equation}
The matrices $\mN$ and $\mM$ contain all the dynamics and energetic properties of the system.
The advantage is that one can now handle simple matrices rather than operators acting on an enormous Hilbert space.
The size of these matrices depends entirely on the set of operators.

@@

We create our operator set on the basis of \eqref{eqn:equal_expecs}.
It consists of operators of the type
\begin{equation}
    A_\gamma \coloneqq \frac{1}{\sqrt{N}} \sum_{\vk} \delta (\gamma - \hat{\gamma}( \vk )) A_{\vk}\,,
\end{equation}
where $A_{\vk}$ represents each type of operator in \eqref{eqn:operators} and their Hermitian conjugates.
Operators with different indices are put into the set individually, e.g., $n_{\gamma \up}$, and $n_{\gamma \down}$ are distinct operators in our set.

Naturally, we have to discretize $\gamma$, 
however, this allows us to obtain good results with drastically smaller matrices compared to a momentum-based operator set and obtain results for a three-dimensional system in the first place.
Practically, we choose $N_\gamma = 3000$ equally spaced discretization points.
A detailed explanation of the numerics is provided in the appendix \textcolor{red}{!REF!}.

Next, we order these operators in a way reminiscent of the $x$- and $p$-terms in the harmonic oscillator, i.e.,
\begin{equation}
    X_i \coloneqq  A_i + A_i^\dagger\,,\quad P_i \coloneqq  A_i - A_i^\dagger\,.
\end{equation}
Then, our set is given by $\mathcal{B}_{XP} \coloneqq \{ X_i, P_i \}$.
If some operators would be 0 ($P_i$ for any $n$-type term) or duplicates (certain $X_i$ for $g$-type terms), we omit them.

Due to this choice of set, the matrices occurring in \eqref{eqn:heisenberg} have a special block structure
\begin{equation}
    \label{eqn:xp_set}
    \mM = \begin{pmatrix}
        \mathcal{K}_+ & \kappa \\ \kappa^\dagger & \mathcal{K}_-
    \end{pmatrix}\,,\quad \mN = \begin{pmatrix}
        \Lambda_+ & \mathcal{L} \\ \mathcal{L}^\dagger & \Lambda_-
    \end{pmatrix}\,,
\end{equation}
where the upper block refers to all $X_i$ and the lower block to all $P_i$.
Furthermore, due to symmetry, the relations $\Im [\mathcal{K}_\pm] = \Re [\kappa] = 0$ and $\Re [\Lambda_\pm] = \Im [\mathcal{L}] = 0$ hold.
Thus, if all occurring expectation values are real, both matrices have large empty blocks.
This condition is fulfilled for all cases investigated in this article.
Additionally, the matrix $\mM$ is positive semidefinite if the system is in thermal equilibrium.
We proof this statement in the appendix \textcolor{red}{!REF!}.

To study the collective excitations, we investigate various retarded Green's functions 
\begin{equation}
    G_{AB}^\text{ret} (t) = - i \langle [A(t), B] \rangle \Theta(t)\,,
\end{equation}
and their Fourier-transformed
\begin{equation}
    \label{eqn:standard_gf}
    G_{AB}(\omega + i0^+) = -i \int_0^{\infty} e^{izt} \langle [A(t), B] \rangle \mathrm{d}t\,,
\end{equation}
where $\Theta(t)$ is the Heaviside function.
Our previous choice of pseudo scalar product enables us to write down a matrix-valued Green's functions in terms of the aforementioned matrices
\begin{align}
    \label{eqn:green_function}
    \mathcal{G}(z &= \omega + i0^+) = \mN \frac{1}{-z \mN - \mM} \mN \\
        &\coloneqq  -\mN \mathcal{R}(z) \mN\,,
\end{align}
where we used the resolvent
\begin{equation}
    \label{eqn:resolvent}
    \mathcal{R}(z) \coloneqq  \frac{1}{z \mN + \mM}\,.
\end{equation}
The entries of this matrix $\mathcal{G}(z)$ are the Fourier-transformed Green's functions mentioned in \eqref{eqn:standard_gf} with respect to the operators from the initially chosen set.
For instance, assume the operator in the set is $f_{\vk}$, then the top left element of $\mathcal{G}$ is the Green's function 
\begin{align}
    \mathcal{G}_{00}(z = \omega +i0^+) &=  G_{f_{\vk}(t) f_{\vk}^\dagger} (z) \nonumber \\
        &= -i \int_0^{\infty} \langle [f_{\vk}(t), f_{\vk}^\dagger(0)] \rangle e^{izt} \mathrm{d}t\,.
\end{align}

In the following, we will derive this relation.
For that, consider the Fourier-transformed Green's function \eqref{eqn:standard_gf} with respect to two operators $A$ and $B^\dagger$ that are included in our operator set
\begin{align}
    G_{AB^\dagger} (z) &= - i \int_0^\infty \exp(izt) \langle [ A(t), B^\dagger(0) ] \rangle \mathrm{d}t
\end{align}
Since $A$ and $B$ are included in the operator set, we can rewrite them as dot products of the vector
\begin{equation}
    \vec{A} \coloneqq  \left( A_1, A_2, \dots, A_N \right)\,, \quad
    \vec{A}^\dagger \coloneqq  \begin{pmatrix}
        A_1^\dagger \\ A_2^\dagger \\ \vdots \\ A_N^\dagger
    \end{pmatrix}\,,
\end{equation}
and some other complex-valued vector $\vec{a}$ which has its coefficients set in such a way,
that $\vec{a} \cdot \vec{A}$ is equal to the desired operator. 
For instance, $\vec{a} \equiv \vec{e}_1$ in order to represent $A_1$.

Furthermore, assuming, that $\mN$ is invertible, we can easily solve the differential equation \eqref{eqn:heisenberg} to obtain
\begin{equation}
    \vec{c}(t) = \exp \left( i \mN^{-1} \mM t \right) \vec{c}(0)\,.
\end{equation}
This allows us to finally obtain
\begin{align}
    G_{AB^\dagger} (z) &= i \int_0^\infty \exp(izt) \langle [ (\vec{A \cdot} \vec{b}(0))^\dagger, \vec{A} \cdot \vec{a}(t) ] \rangle\ \mathrm{d}t \nonumber \\
        &= i \int_0^\infty \exp(izt) \vec{b}^\dagger(0) \underbrace{\langle [ \vec{A}^\dagger, \vec{A} ] \rangle}_{\equiv \mN} \vec{a}(t) \mathrm{d}t \nonumber \\
        &= i  \int_0^\infty \vec{b}^\dagger(0) \mN \exp \left( i \left(\mN^{-1} \mM + z \right) t \right) \vec{a}(0) \mathrm{d}t  \nonumber \\
        &= - \vec{b}^\dagger(0) \left[ \mN \frac{1}{\mN^{-1} \mM + z} \right] \vec{a}(0) \nonumber \\
        &= - \vec{b}^\dagger(0) \left[ \mN \frac{1}{ \underbrace{\mM + z \mN}}_{\equiv \mathcal{R}(z)} \mN \right] \vec{a}(0)\,,
\end{align}
where $\vec{a}(0)$ and $\vec{b}(0)$ are chosen such, that their product with $\vec{A}$ yields $A(0)$ and $B(0)$, respectively.
The task at hand is finding the resolvent as $\mathcal{R}$.

To do that, we embrace this matrix structure in \eqref{eqn:xp_set} and assume that all matrix entries are real.
Then, with $\mathcal{R}|_X$ denoting the block of a matrix $\mathcal{R}$ corresponding to the $X$-operators, we can calculate
\begin{align}
    r(z) &\coloneqq  \mathcal{R}(z)|_X = \left. \frac{1}{\mM + z \mN} \right\vert_X \nonumber \\
        &= \left[ \frac{1}{\mM} - \frac{1}{\mM} z \mN \frac{1}{\mM} + \frac{1}{\mM} z \mN \frac{1}{\mM} z \mN \frac{1}{\mM} - \cdots \right]_X \nonumber \\
        &= \left. \frac{1}{\mM} \sum_{j=0}^\infty \left( -z \mN \frac{1}{\mM} \right)^j \right\vert_X \,.
\end{align}
We may omit every second element of the sum, as we want to obtain the upper left block of the matrix, and $\mN$ always swaps the upper left and lower right block.
Thus, we obtain
\begin{align}
    r(z) &= \left. \frac{1}{\mM} \sum_{j=0}^\infty z^{2j} \left( \mN \frac{1}{\mM} \right)^{2j} \right\vert_X \nonumber \\
        &= \left. \frac{1}{\mM - z^2 \mN \mM^{-1} \mN} \right\vert_X \nonumber \\
        &= \frac{1}{\mathcal{K}_+ - z^2 \mathcal{L} \mathcal{K}_-^{-1} \mathcal{L}^\dagger}\,.
\end{align}
Next, we define $\check{N} \coloneqq \mathcal{L} \mathcal{K}_-^{-1} \mathcal{L}^\dagger$ for brevity.
Since $\mathcal{M}$ is positive semidefinite and block diagonal, its blocks $\mathcal{K}_\pm$ must be positive semidefinite as well.
Therefore, $\check{N}$ is also positive semidefinite and we can define its square root.
We obtain
\begin{equation}
    r(z) = \check{N}^{-1/2} \frac{1}{\check{N}^{-1/2} \mathcal{K}_+ \check{N}^{-1/2} - z^2} \check{N}^{-1/2}\,.
\end{equation}
The matrix $\check{M} \coloneqq  \check{N}^{-1/2} \mathcal{K}_+ \check{N}^{-1/2}$ is positive semidefinite and Hermitian by the previous arguments.
If $\check{N}$ has a singular part, the inverse can be replaced by the pseudo-inverse.
The same calculation can be repeated by exchanging $X$ with $P$, yielding the relation for the remaining block.

@@

\begin{widetext}
\begin{equation}
    \left( z^2 - \begin{pmatrix}
        a_0 & b_1 & 0 & 0 & \cdots \\
        b_1 & a_1 & b_2 & 0 & \cdots \\
        0 & b_2 & a_2 & b_3 & \cdots \\
        0 & 0 & b_3 & a_3 & \ddots \\
        \vdots & \vdots & \vdots & \ddots & \ddots
    \end{pmatrix} \right)_{00}^{-1} = \dfrac{1}{z^2 - a_0 - \dfrac{b_1^2}{z^2 - a_1 - \dfrac{b_2^2}{ z^2 - a_2 - \hdots}}}\,\,
\end{equation}
\end{widetext}

where $a_i$ and $b_i$ are the Lanczos coefficients \cite{PettiforRecursion,ViswanathRecursion}.
In the single-band case, the coefficients approach the limit
\begin{equation}
    \label{eqn:inf_lanczos}
    a_\infty = \frac{\omega_+ + \omega_-}{2}\text{  and  } b_\infty = \frac{\omega_+ - \omega_-}{4}\,,
\end{equation}
where $\omega_\pm$ represents the upper and lower band edge, respectively \cite{PettiforRecursion}.
This behaviour allows us to terminate the continued fraction, once the coefficients have converged close enough to these values, using the so-called square root terminator
\begin{equation}
    T(\omega) = \frac{1}{2b_\infty^2} \left( \omega - a_\infty \mp \sqrt{(\omega - a_\infty)^2 - 4 b_\infty^2} \right)\,,
\end{equation}
where the negative sign is to be chosen for $\omega - a_\infty > 2b_\infty$ and $|\omega - a_\infty| < 2b_\infty$ and the positive one otherwise.
It is to be noted, that the coefficients start to drop off from this limit as more and more Lanczos iterations are performed \cite{ViswanathRecursion}.
Therefore, we terminate the continued fraction with the aforementioned terminator at the coefficient pair that has the least deviation from $a_\infty$ and $b_\infty$.

@@

In this section, we will be investigating four different diagonal Green's functions $G_{AA^\dagger}(\omega + i0^+)$, namely with
\begin{subequations}
    \label{eqn:resolvent_bases}
    \begin{align}
        A_\text{Higgs} &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( f_{\vk} + f_{\vk}^\dagger \right) \nonumber \\ 
            &= \int_{-1}^{1} \rho(\gamma) \left( f(\gamma) + f^\dagger (\gamma) \right) \mathrm{d}\gamma \,,\\
        A_\text{Phase} &= \frac{i}{\sqrt{N}} \sum_{\vk} \left( f_{\vk} - f_{\vk}^\dagger \right) \nonumber \\ 
            &= i \int_{-1}^{1} \rho(\gamma) \left( f(\gamma) - f^\dagger (\gamma) \right) \mathrm{d}\gamma \,,\\
        A_\text{CDW} &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( g_{\vk \up} + g_{\vk \down} \right) \nonumber \\ 
            &= \int_{-1}^{1} \rho(\gamma) \left( g_{\up}(\gamma) + g_{\down}(\gamma) \right) \mathrm{d}\gamma \,,\\
        A_\text{AFM} &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( g_{\vk \up} - g_{\vk \down} \right) \nonumber \\ 
            &= \int_{-1}^{1} \rho(\gamma) \left( g_{\up}(\gamma) - g_{\down}(\gamma) \right) \mathrm{d}\gamma \,,
    \end{align}
\end{subequations}
where $N$ is the number of lattice sites in the system. 
Each one of these operators generates a different kind of collective mode.
The first operator captures the amplitude mode of the $s$-wave superconducting state, while the second one catches the phase mode \cite{Fan22}.
The remaining two capture the collective behaviour of the CDW and AFM states, respectively.
Note, that this choice of normalization cancels out with other system size-dependent terms so that our final results do not depend on it at all.

@@

Nevertheless, let us show that $\mathcal{M}$ is positive semidefinite if the system is in thermal equilibrium.
For any $\vec{x} \in \mathbb{C}^N$, consider
\begin{align}
    \vec{x}^\dagger \mM \vec{x} &= \sum_{ij} x_i^* x_j \mM_{ij} \nonumber \\
        &= \langle \left[ \sum_i x_i^* A_i^\dagger, \left[ H, \sum_j x_j A_j \right] \right]  \rangle \nonumber \\
        &= \langle [B^\dagger, [H, B]] \rangle\,,\quad B \coloneqq  \sum_j x_j A_j\,.
\end{align}
Expanding the thermal expectation value $\langle O \rangle = \frac{1}{Z} \Trace [\exp(-\beta H_\text{MF}) O]$, where $\beta$ is the inverse temperature and $Z$ the partition function, and the commutators, we obtain

@@
